---
title: "BVS.DAdj quick user guidance"
author: "Yubing Yao"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This page gives a step-by-step workflow for using the **BVS.DAdj** R package
for Bayesian variable selection with dual-adjacency network-informed priors.
The workflow is fully in R. It does **not** require MATLAB code or external
MATLAB sampling of Bayesian Gaussian graphical models for adjacency estimation.

In the example below, the two adjacency sources are:

1. an external/database adjacency matrix, for example from KEGG, HMDB,
   Reactome, pathway annotations, or prior biological knowledge;
2. a data-driven adjacency matrix estimated from the observed predictor matrix
   with graphical lasso in R.

For binary outcomes, use `bvs_pg()` or `bvs_mh()`. For continuous outcomes, use
`bvs_mh(outcome_type = "continuous")`.

## Install and load the package

```{r install-load}
# Install from GitHub if needed.
# install.packages("devtools")
# devtools::install_github("yuy113/BVS.DAdj")

# Suggested packages for this guide.
# install.packages(c("MASS", "huge", "coda"))

library(BVS.DAdj)
library(MASS)
```

The package function `estimate_glasso_adj()` uses the `huge` package to build a
binary adjacency matrix from graphical lasso. Install `huge` when using
`adj_type = "glasso"` or when calling `estimate_glasso_adj()` directly.

## Prepare outcome, predictors, and optional covariates

The predictor matrix `X` should have observations in rows and predictors in
columns. Standardizing the predictor columns is recommended before estimating a
data-driven adjacency matrix.

```{r simulate-data}
set.seed(2026)

n <- 120
p <- 40
active <- 1:6

# Correlated active block plus noise predictors.
Sigma <- diag(p)
Sigma[active, active] <- 0.35
diag(Sigma) <- 1
X <- MASS::mvrnorm(n = n, mu = rep(0, p), Sigma = Sigma)
X <- scale(X)
colnames(X) <- paste0("M", seq_len(p))

# Optional clinical covariate not under variable selection.
Z <- matrix(rnorm(n), ncol = 1)
colnames(Z) <- "clinical_covariate"

# Binary outcome.
beta_true <- rep(0, p)
beta_true[active] <- c(1.0, 1.0, 0.8, 0.7, 0.6, 0.5)
eta <- -0.4 + X %*% beta_true + 0.8 * Z[, 1]
prob <- plogis(eta)
y_binary <- rbinom(n, size = 1, prob = prob)

# Continuous outcome, if needed.
y_continuous <- as.numeric(0.5 + X %*% beta_true + 0.8 * Z[, 1] +
                             rnorm(n, sd = 1))
```

For real data, replace `X`, `Z`, and `y_binary` or `y_continuous` with your
cleaned analysis data.

## Build an external/database adjacency matrix

The external adjacency matrix must be a square `p x p` matrix with entries 0 or
1, zero diagonal, and the same predictor order as the columns of `X`.

In a real analysis, construct this matrix from a database-derived edge list
such as KEGG, HMDB, Reactome, taxonomy, or pathway co-membership. The example
below mimics pathway co-membership.

```{r external-adjacency}
# Example pathway labels. Replace this object with real database annotations.
pathway_label <- rep("none", p)
pathway_label[1:10] <- "pathway_A"
pathway_label[11:18] <- "pathway_B"

R_external <- outer(pathway_label, pathway_label, FUN = "==") * 1L
R_external[pathway_label == "none", ] <- 0L
R_external[, pathway_label == "none"] <- 0L
diag(R_external) <- 0L
rownames(R_external) <- colnames(R_external) <- colnames(X)

sum(R_external) / 2
```

If you already have an edge list, convert it to an adjacency matrix:

```{r external-edge-list}
edge_list <- data.frame(
  node1 = c("M1", "M2", "M4"),
  node2 = c("M2", "M3", "M5")
)

R_external_from_edges <- matrix(0L, p, p,
                                dimnames = list(colnames(X), colnames(X)))
idx1 <- match(edge_list$node1, colnames(X))
idx2 <- match(edge_list$node2, colnames(X))
ok <- !is.na(idx1) & !is.na(idx2) & idx1 != idx2
R_external_from_edges[cbind(idx1[ok], idx2[ok])] <- 1L
R_external_from_edges[cbind(idx2[ok], idx1[ok])] <- 1L
diag(R_external_from_edges) <- 0L
```

## Build a data-driven graphical-lasso adjacency matrix in R

No MATLAB step is needed. The observed predictor matrix `X` can be passed to
`estimate_glasso_adj()`:

```{r glasso-adjacency}
R_glasso <- estimate_glasso_adj(
  X,
  criterion = "ebic",
  nlambda = 30,
  lambda.min.ratio = 0.01
)

rownames(R_glasso) <- colnames(R_glasso) <- colnames(X)
sum(R_glasso) / 2
```

This matrix can be used as the second adjacency source in a dual-adjacency
model. You can also let the package estimate graphical lasso internally by
using `adj_type = "glasso"` or `adj_type = "glasso_fixed"`, but explicitly
calling `estimate_glasso_adj()` makes the workflow easier to inspect and
reproduce.

## Run BVS.DAdj for a binary outcome

For binary logistic outcomes, `bvs_pg()` uses a Polya-Gamma Gibbs sampler.
Here the external adjacency and the graphical-lasso adjacency are supplied as
two fixed adjacency sources through `adj_type = "dual_fixed"`.

```{r binary-bvs-pg}
fit_pg_dual <- bvs_pg(
  X = X,
  y = y_binary,
  z_dat = Z,
  adj_type = "dual_fixed",
  adj_fixed = R_external,
  adj_fixed2 = R_glasso,
  niter = 60000,
  burnin = 10000,
  thin = 5,
  eta1_sd = 0.5,
  eta2_sd = 0.5,
  store_beta = TRUE,
  store_gamma = TRUE
)
```

The same binary outcome can also be fit with the Metropolis-Hastings sampler:

```{r binary-bvs-mh}
fit_mh_binary <- bvs_mh(
  X = X,
  y = y_binary,
  outcome_type = "binary",
  z_dat = Z,
  adj_type = "dual_fixed",
  adj_fixed = R_external,
  adj_fixed2 = R_glasso,
  niter = 60000,
  burnin = 10000,
  thin = 5,
  eta1_sd = 0.5,
  eta2_sd = 0.5,
  store_beta = TRUE,
  store_gamma = TRUE
)
```

A shorter equivalent workflow lets the package estimate the graphical-lasso
adjacency internally:

```{r binary-glasso-fixed-shortcut}
fit_pg_glasso_fixed <- bvs_pg(
  X = X,
  y = y_binary,
  z_dat = Z,
  adj_type = "glasso_fixed",
  adj_fixed = R_external,
  glasso_criterion = "ebic",
  niter = 60000,
  burnin = 10000,
  thin = 5
)
```

## Run BVS.DAdj for a continuous outcome

For continuous Gaussian outcomes, use `bvs_mh()` with
`outcome_type = "continuous"`.

```{r continuous-bvs-mh}
fit_mh_continuous <- bvs_mh(
  X = X,
  y = y_continuous,
  outcome_type = "continuous",
  z_dat = Z,
  adj_type = "dual_fixed",
  adj_fixed = R_external,
  adj_fixed2 = R_glasso,
  niter = 60000,
  burnin = 10000,
  thin = 5,
  eta1_sd = 0.5,
  eta2_sd = 0.5,
  h_beta = 25,
  store_beta = TRUE,
  store_gamma = TRUE
)
```

The internal graphical-lasso shortcut is also available for continuous
outcomes:

```{r continuous-glasso-fixed-shortcut}
fit_mh_cont_glasso_fixed <- bvs_mh(
  X = X,
  y = y_continuous,
  outcome_type = "continuous",
  z_dat = Z,
  adj_type = "glasso_fixed",
  adj_fixed = R_external,
  glasso_criterion = "ebic",
  niter = 60000,
  burnin = 10000,
  thin = 5,
  h_beta = 25
)
```

## Optional: Bayesian GGM adjacency inside BVS.DAdj

If you want Bayesian Gaussian graphical model adjacency learning rather than
graphical lasso, use `adj_type = "ggm"` or `adj_type = "ggm_fixed"` directly
in `bvs_mh()` or `bvs_pg()`. This is handled by the **BVS.DAdj** R package and
does not require MATLAB scripts or exported adjacency text files.

```{r internal-ggm-note}
fit_internal_ggm <- bvs_mh(
  X = X,
  y = y_binary,
  outcome_type = "binary",
  z_dat = Z,
  adj_type = "ggm_fixed",
  adj_fixed = R_external,
  niter = 60000,
  burnin = 10000,
  thin = 5,
  store_Z_pip = TRUE
)
```

For high-dimensional data, set `sparse = TRUE` for `adj_type = "ggm"` or
`adj_type = "ggm_fixed"`. If desired, precompute the sparse scatter matrix with
`prepare_sparse_S(X)`.

## Summarize posterior inclusion probabilities

Posterior inclusion probabilities (PIPs) are used to rank and select
predictors. A common initial threshold is 0.5, but the threshold should be
chosen for the scientific goal and the desired false-positive/false-negative
tradeoff.

```{r posterior-summary}
summary(fit_pg_dual)

pip <- if (!is.null(fit_pg_dual$gamma_pip)) {
  fit_pg_dual$gamma_pip
} else {
  colMeans(fit_pg_dual$gamma)
}

pip_table <- data.frame(
  predictor = colnames(X),
  PIP = as.numeric(pip)
)
pip_table <- pip_table[order(pip_table$PIP, decreasing = TRUE), ]
head(pip_table, 10)

selected <- subset(pip_table, PIP >= 0.5)
selected
```

## Diagnostic plots

The package provides PIP plots, trace plots, and ACF plots for fitted BVS
objects.

```{r diagnostics}
plot(fit_pg_dual, type = "pip", pip_threshold = 0.5)
plot(fit_pg_dual, type = "trace", top_n = 6)
plot(fit_pg_dual, type = "acf", top_n = 6)
```

For multiple chains, compare PIPs and selected predictors across independent
seeds. The package also includes convergence and selection diagnostic utilities
for fitted objects and posterior samples.

## Recommended practical workflow

1. Clean and standardize `X`; align all sample IDs and predictor names.
2. Build `R_external` from a trusted external database or annotation source.
3. Build `R_glasso` in R with `estimate_glasso_adj(X, criterion = "ebic")`.
4. Fit a dual-adjacency model with `adj_type = "dual_fixed"`, or use
   `adj_type = "glasso_fixed"` to estimate graphical lasso inside the package.
5. Run enough MCMC iterations for the data size and check trace/ACF diagnostics.
6. Rank predictors by PIP and report selected predictors with the chosen PIP
   threshold.

## Contact and links

The R package is available at
<https://github.com/yuy113/BVS.DAdj>.

For questions, please contact Yubing Yao.

## References

- Polson, N. G., Scott, J. G., and Windle, J. (2013). Bayesian inference for
  logistic models using Polya-Gamma latent variables. *Journal of the American
  Statistical Association*.
- Friedman, J., Hastie, T., and Tibshirani, R. (2008). Sparse inverse
  covariance estimation with the graphical lasso. *Biostatistics*.
- Zhao, T., Liu, H., Roeder, K., Lafferty, J., and Wasserman, L. (2012). The
  huge package for high-dimensional undirected graph estimation in R. *Journal
  of Machine Learning Research*.
