---
title: "Active subnetwork detection Quick Start"
author: "Yubing Yao"
date: '2023-05-17'
output: 
  rmdformats::robobook:
    code_folding: show
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install and Load the `wActNet` Package


Begin by installing and loading the `devtools` package, then using the `install_github` function to install wActNet from GitHub as follows and we also need additional 2 R packages of `igraph` and `BioNet` from BioConductor <https://www.bioconductor.org/install/>.



```{r,results='hide' }
#please comment the below code out if you already install the package-wActNet from github
#library(devtools)
#if (!requireNamespace("BiocManager", quietly = TRUE))
#  install.packages("BiocManager")
#BiocManager::install("pathview")
#install_github("yuy113/wActNet")

library(wActNet)
#R package BioNet from BioConductor 
library(BioNet)
library(igraph)
```

## Simulate data and generate P values of tests on pairwise correlations

Next, for illustration, we generate pairwise correlation data from simulating random data based on multivariate normal distribution, and we could perform multiple nonparametric permutation tests for all pairwise correlations of simulated data and get corresponding P values from the permutation tests through the function `pval.perm.corr`. 


```{r}
dat1<-matrix(rnorm(4000),ncol=40,nrow=100)
colnames(dat1)<-paste("Var", as.character(1:40),sep="")
pval.edge<-pval.perm.corr(dat1,nsim=100,MatchId=NULL,do.parallel=FALSE)
#pval.edge
```

## Calculate node scores and edge scores 


Simulate p values for all nodes and all possible connecting edges in one illustrated network with 40 nodes, through the function `uniform.beta.node.edge.score` to generate the list output of node score-`NodeScore`, edge scores-`EdgeScore` and network data of `igraph object`-`Network`.


```{r}
# simulate the p values for all the possible edges in the network
ind.pos.pval.edge<-rbinom(40*39/2,1,0.5)
pval.edge<-(1-ind.pos.pval.edge)*runif(40*39/2)+ind.pos.pval.edge*rbeta(40*39/2,0.1,1)
names(pval.edge)<-unlist(sapply(1:39,function(i){sapply((i+1):40, function(j){paste(paste("Var",
as.character(i),sep=""),paste("Var",as.character(j),sep=""),sep="_")})}))

# simulate p values for all the nodes in the network
ind.pos.pval.node<-rbinom(40,1,0.2)
pval.node<-(1-ind.pos.pval.node)*runif(40)+ind.pos.pval.node*rbeta(40,0.1,1)
names(pval.node)<-paste("Var", as.character(1:40),sep="")

# generate the node score-NodeScore, edge scores-EdgeScore and igraph object-Network
network.test<-uniform.beta.node.edge.score(pval.node,pval.edge,0.05,0.05,dat1)

#igraph object-Network
network.test$Network
#node scores
network.test$NodeScore
#edge scores
#network.test$EdgeScore
```


## Detect optimized subnetworks 


run an example to detect all possible optimized subnetworks on simulated network data above, we could run the method of Dittrich et al. 2008 based on node scores only and our proposed method based on weighted combined node scores and edge scores respectively throught the function `MultiModuleFind`. In the function `MultiModuleFind` we could set up the maximum possible number of acctive subnetworks-`ncluster` and select 2 methods of Dittrich et al. 2008 or our proposed method in `method` parameter of `NodeOnly` and `NodeEdge` respectively. Also in our proposed method implemented in `wActNet` R package we could also adjust the parameter-`weightratio.edge.node`: weighted ratio of edge scores versus node scores in objective function for optimizing active subnetworks.

```{r}
network.test1<-network.test$Network
node.scores<-network.test$NodeScore
edge.scores<-network.test$EdgeScore

#identify all possible optimized subnetworks in simulated network
# using Dittrich's method using node scores only
multi.mod.n<-MultiModuleFind(network.test1,node.scores,edge.scores,
weightratio.edge.node=1,ncluster=3,method="NodeOnly")
#all possible optimized subnetworks in simulated network using Dittrich's method using node scores only
multi.mod.n
# identify all possible signaling modules in simulated network
# using our proposed method based on node scores and edge scores
multi.mod.e<-MultiModuleFind(network.test1,node.scores,edge.scores,
weightratio.edge.node=1,ncluster=3,method="NodeEdge")
#all possible optimized subnetworks in simulated network using our proposed method based on node scores and edge scores
multi.mod.e
```


## Contact Us / Contribute
This package is new, and any and all suggestions are welcomed. You can use GitHub to raise issues, contribute, or communicate with us about the package:

https://github.com/yuy113/wActNet/

## Reference

Dittrich, M., Klau, G., Rosenwald, A., Dandekar, T., Müller, T.: Identifying functional modules in protein-protein interaction
networks: an integrated exact approach. Bioinformatics 24(13), 223–231 (2008)

Beisser, D., Klau, G., Dandekar, T., Müller, T., Dittrich, M.: Bionet: an r-package for the functional analysis of biological
networks. Bioinformatics 26(8), 1129–1130 (2010)

