---
title: "Active subnetwork detection Quick Start"
author: "Yubing Yao"
date: '2023-05-17'
output: 
  rmdformats::robobook:
    code_folding: show
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install and Load the `wActNet` Package

Begin by installing and loading the `devtools` package. Use the install_github function to install `wActNet` from `GitHub` as shown below. Our package depends on two additional R packages, `igraph` and `BioNet`, and the second one can be downloaded from `BioConductor` <https://www.bioconductor.org/install/>.


```{r,results='hide' }
#please comment the below code out if you already install the package-wActNet from github
#library(devtools)
#if (!requireNamespace("BiocManager", quietly = TRUE))
#  install.packages("BiocManager")
#BiocManager::install("pathview")
#install_github("yuy113/wActNet")

library(wActNet)
#R package BioNet from BioConductor 
library(BioNet)
library(igraph)
```

## Simulate data and generate P values of tests on pairwise correlations


To illustrate the process, we simulate data ($n$ observations, $p$ features) from a multivariate normal distribution, which is then used to generate data on ${p}\choose{2}$ pairwise correlations data through the simulation of random data, which follows a multivariate normal distribution. Subsequently, we use the function `pval.perm.corr()` to conduct multiple conduct nonparametric permutation tests for all the pairwise correlations within the simulated data. Using the function and obtaining corresponding P values, we obtain the corresponding P-values from these permutation tests.



```{r}
dat1<-matrix(rnorm(4000),ncol=40,nrow=100)
colnames(dat1)<-paste("Var", as.character(1:40),sep="")
pval.edge<-pval.perm.corr(dat1,nsim=100,MatchId=NULL,do.parallel=FALSE)
#pval.edge
```

## Calculate node scores and edge scores 


To illustrate our approach, we begin by simulating a network of 40 nodes. We generate P values for all nodes and potential connecting edges using the `uniform.beta.node.edge.score()` function. The output of this function is a list consisting of node scores (`NodeScore`), edge scores (`EdgeScore`), and the network data in the form of an igraph object (`Network`).



```{r}
# simulate the p values for all the possible edges in the network
ind.pos.pval.edge<-rbinom(40*39/2,1,0.5)
pval.edge<-(1-ind.pos.pval.edge)*runif(40*39/2)+ind.pos.pval.edge*rbeta(40*39/2,0.1,1)
names(pval.edge)<-unlist(sapply(1:39,function(i){sapply((i+1):40, function(j){paste(paste("Var",
as.character(i),sep=""),paste("Var",as.character(j),sep=""),sep="_")})}))

# simulate p values for all the nodes in the network
ind.pos.pval.node<-rbinom(40,1,0.2)
pval.node<-(1-ind.pos.pval.node)*runif(40)+ind.pos.pval.node*rbeta(40,0.1,1)
names(pval.node)<-paste("Var", as.character(1:40),sep="")

# generate the node score-NodeScore, edge scores-EdgeScore and igraph object-Network
network.test<-uniform.beta.node.edge.score(pval.node,pval.edge,0.05,0.05,dat1)

#igraph object-Network
network.test$Network
#node scores
network.test$NodeScore
#edge scores
#network.test$EdgeScore
```


## Detect optimized subnetworks 


To detect active subnetworks within our network, we illustrate the approach by Dittrich et al. in 2008 and compare to the `wActNet` algorithm. The algorithm proposed by Dittrich et al. in 2008 uses node scores only. The `wActNet` algorithm combines node scores and edge scores using the `MultiModuleFind` function. Within the `MultiModuleFind` function, we have the flexibility to set the maximum number of active subnetworks (`ncluster`) and choose between two methods: the one presented by Dittrich et al. in 2008 (method=`NodeOnly`) or the wActNet approach (method=`NodeEdge`). 

Furthermore, the wActNet algorithm allows us to fine-tune the optimization of active subnetworks by adjusting the parameter `weightratio.edge.node`. This parameter represents the weighted ratio between edge scores and node scores in the objective function.

In summary, by combining the methodology introduced by Dittrich et al. (2008) with the `wActNet` extension that includes both  node and edge scores, we can effectively detect the optimized active subnetworks on network data.


Please note when the function completes execution and fails to identify any further optimized or active subnetworks, a warning message "No positive nodes" will be generated.


```{r}
network.test1<-network.test$Network
node.scores<-network.test$NodeScore
edge.scores<-network.test$EdgeScore

#Note: both BioNet and wActNet R packages stop running and can't detect any optimized/active subnetwork further and complete, a warning output message-"No positive nodes" will be generated.
#identify all possible optimized subnetworks in simulated network
# using Dittrich's method using node scores only
multi.mod.n<-MultiModuleFind(network.test1,node.scores,edge.scores,
weightratio.edge.node=1,ncluster=3,method="NodeOnly")
#all possible optimized subnetworks in simulated network using Dittrich's method using node scores only
multi.mod.n
#visualize the final illustrated detected optimized subnetwork by BioNet algorithm
#for edges, edge score >=0, solid line, edge score < 0, dash line
#for nodes, node score >=0, red color, edge score < 0, grey color 
#the size/length of node or edges proportional to node or edge score respectively
#need R package-visNetwork, if not, please install from CRAN
require(visNetwork, quietly = TRUE)
dat_node<-multi.mod.n[[1]]
 E(dat_node)$dashes<-ifelse(E(dat_node)$score>0,FALSE,FALSE)
  E(dat_node)$width<-(E(dat_node)$score+1.5)*0.3
  V(dat_node)$color<-ifelse(V(dat_node)$score>0,"#E7298A","grey")
  V(dat_node)$size<-(V(dat_node)$score+11)*0.5
  
  E(dat_node)$color<-ifelse(E(dat_node)$score>0,"#66A61E","grey")
  
  
 #########################################################################################
  # setwd("/Users/yubingyao/Google Drive/Network analysis/R code/")
  
 # save(GU2,file="network2_2clusters_plot.RData")
  
  
  
  #load("network_2clusters_plot.RData")
  
  data_node<- toVisNetworkData(dat_node)
  
  visNetwork(nodes = data_node$nodes, edges = data_node$edges, height = "800px")%>% visNodes(font=list(size=20))
# identify all possible signaling modules in simulated network
# using our proposed method based on node scores and edge scores
multi.mod.e<-MultiModuleFind(network.test1,node.scores,edge.scores,
weightratio.edge.node=1,ncluster=3,method="NodeEdge")
#all possible optimized subnetworks in simulated network using our proposed method based on node scores and edge scores
multi.mod.e

#visualize the final illustrated detected optimized subnetwork by our proposed algorithm
#for edges, edge score >=0, solid line, edge score < 0, dash line
#for nodes, node score >=0, red color, edge score < 0, grey color 
#the size/length of node or edges proportional to node or edge score respectively
#need R package-visNetwork, if not, please install from CRAN
#require(visNetwork, quietly = TRUE)
dat_nodeedge<-multi.mod.e[[1]]
 E(dat_nodeedge)$dashes<-ifelse(E(dat_nodeedge)$score>0,FALSE,FALSE)
  E(dat_nodeedge)$width<-(E(dat_nodeedge)$score+1.5)*0.3
  V(dat_nodeedge)$color<-ifelse(V(dat_nodeedge)$score>0,"#E7298A","grey")
  V(dat_nodeedge)$size<-(V(dat_nodeedge)$score+11)*0.5
  
  E(dat_nodeedge)$color<-ifelse(E(dat_nodeedge)$score>0,"#66A61E","grey")
  
  
 #########################################################################################
  # setwd("/Users/yubingyao/Google Drive/Network analysis/R code/")
  
 # save(GU2,file="network2_2clusters_plot.RData")
  
  
  
  #load("network_2clusters_plot.RData")
  
  data_nodeedge <- toVisNetworkData(dat_nodeedge)
  
  visNetwork(nodes = data_nodeedge$nodes, edges = data_nodeedge$edges, height = "800px")%>% visNodes(font=list(size=20))
```


## Contact Us / Contribute
This package is new, and any and all suggestions are welcomed. You can use GitHub to raise issues, contribute, or communicate with us about the package:

https://github.com/yuy113/wActNet/

## Reference

Dittrich, M., Klau, G., Rosenwald, A., Dandekar, T., Müller, T.: Identifying functional modules in protein-protein interaction
networks: an integrated exact approach. Bioinformatics 24(13), 223–231 (2008)

Beisser, D., Klau, G., Dandekar, T., Müller, T., Dittrich, M.: Bionet: an r-package for the functional analysis of biological
networks. Bioinformatics 26(8), 1129–1130 (2010)

